---
title: "Data Cleaning_Session 8"
author: "Julia Masterman"
date: "3/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, error=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

#Data Sources
##Read in the data
```{r}
catch_original <- read_csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1"))
# you can do this with the URL since this data set is in the repository so the URL will persist. 
```

#Clean and Reshape the Data 
##remove unnecessary columns
```{r}
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
#Could also do select(Region:Chum)
head(catch_data)
```
##check column typing
One variable (Chinook) is a character class, so we  need to fix that. First chunk: investigate 
```{r}
summary(catch_data)
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))
#NAs introduced by coercion
head(catch_clean)
i <- which(is.na(catch_clean$Chinook))
i
catch_data[i,]
catch_clean[i,]
#dataset if from a PDF which is automatically converted to a CSV and the value of I is actually 1
```
Fixing the issue below
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))
head(catch_clean)
```
##reshape data
Making the long form data frame
```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch")
head(catch_long)

#practice to make wide data frame 
catch_wide <- catch_long %>% 
  pivot_wider(names_from = species, values_from = catch)
head(catch_wide)
```
The measurement of catch is reported in thousands, so we should add something to make that clearer. We rename the column 
```{r}
catch_long <- catch_long %>% 
  rename(catch_thousands = catch)
catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000)
head(catch_long)
```
```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(catch_mean = mean (catch)) %>% 
  arrange(desc(catch_mean))
##using arrange to order the site names in a more meaningful way 
##note automatically will sort in acending order 
head(mean_region)

n_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(n = n())
head(n_region)
```

```{r}
SSE_catch <- catch_long %>% 
  filter(Region == "SSE")
table(SSE_catch$Region)
```

#Join to Regions dataset 
```{r}
region_defs <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```

Example of one thing in every pipe from workshop materials. 
```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),stringsAsFactors = FALSE)
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch") %>%
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  left_join(region_defs, by = c("Region" = "code"))

head(mean_region)
```

